### 2014-12-06 text/Chapter01.tex
Proofreading (Chapter 3)

\begin{itemize}
\item \textbf{System capabilities.} Libraries such as LINQ \cite{app-linq} let developers write code
  in a host language like C\# and then cross-compile it to multiple targets (including SQL, OpenCL 
  or JavaScript \cite{app-hop-lang}). Part of the compilation (\eg~generating the SQL query) occurs 
  at runtime and developers have no guarantee that it will succeed until the program is executed,
  because only subset of the host language is supported.

\item \textbf{Platform versions.} When developing cross-platform applications, different platforms
  (and different versions of the same platform) provide different API functions. Writing a
  cross-platform code usually relies on (fragile) conditional compilation or (equally fragile)
  dynamic loading.

\item \textbf{Security and provenance.} When working with data (be it sensitive database or social
   network data), we may have permission to access only some of the data and we may want to track
   \emph{provenance} information. However, this is not checked -- if a program attempts to access
   unavailable data, the access will be refused at run-time.

\item \textbf{Resources \& data availability.} When creating a mobile application, the program may
  (or may not) be granted access to device capabilities such as GPS sensor, social updates or battery
  status. We would like to know which of the capabilities are required and which are optional
  (\ie~enhance the user experience, but there is a fallback strategy). Equally, on the server-side,
  we might have access to different database tables, depending on the role of the user.
\end{itemize}

@@

\begin{itemize}
\item \textbf{Flat coeffects} represent additional data, resources and meta-data that are
   available in the execution environment (regardless of how they are accessed in a program).
   Examples include resources such as GPS sensors and battery status (on a phone), databases
   (on the server), or software framework (or library) version.

\item \textbf{Structural coeffects} capture additional meta-data related to inputs. This can include
   provenance (source of the input value), usage information (how often is the value
   accessed and in what ways) or security information (whether it contain sensitive data or not).
\end{itemize}
%
This thesis follows the tradition of statically typed programming languages. As such, we
attempt to capture such contextual information in the type system of context-aware programming
languages. The type system should provide both safety guarantees (as in the first three examples)
and also static analysis useful for optimization (as in the last example).

@@

\begin{itemize}
\item In Chapter~\ref{ch:pathways} we give an overview of related work. Most importantly, we show
  that the idea of context-aware computations can be naturally approached from a number of
  directions developed recently in the theory of programming languages (including type and effect
  systems, categorical semantics and sub-structural logics).

\item Chapter~\ref{ch:applications} presents the first contribution of the thesis -- the discovery
  of the connection between a number of existing programming language features that are related
  to context. The chapter presents type systems and semantics for a number of systems and analyses 
  (including data-flow, liveness analysis, distributed programming and Haskell's type classes).
  Our novel presentation reveals their similarity.

>>>>

\begin{itemize}
\item \textsc{System capabilities.} Libraries such as LINQ \cite{app-linq} let developers write code
  in a host language like C\# and then cross-compile it to multiple targets (including SQL, OpenCL 
  or JavaScript \cite{app-hop-lang}). Part of the compilation (\eg~generating the SQL query) occurs 
  at runtime and developers have no guarantee that it will succeed until the program is executed,
  because only subset of the host language is supported.

\item \textsc{Platform versions.} When developing cross-platform applications, different platforms
  (and different versions of the same platform) provide different API functions. Writing a
  cross-platform code usually relies on (fragile) conditional compilation or (equally fragile)
  dynamic loading.

\item \textsc{Security and provenance.} When working with data (be it sensitive database or social
   network data), we may have permission to access only some of the data and we may want to track
   \emph{provenance} information. However, this is not checked -- if a program attempts to access
   unavailable data, the access will be refused at run-time.

\item \textsc{Resources \& data availability.} When creating a mobile application, the program may
  (or may not) be granted access to device capabilities such as GPS sensor, social updates or battery
  status. We would like to know which of the capabilities are required and which are optional
  (\ie~enhance the user experience, but there is a fallback strategy). Equally, on the server-side,
  we might have access to different database tables, depending on the role of the user.
\end{itemize}

@@

\begin{itemize}
\item \textsc{Flat coeffects} -- represent additional data, resources and meta-data that are
   available in the execution environment (regardless of how they are accessed in a program).
   Examples include resources such as GPS sensors and battery status (on a phone), databases
   (on the server), or software framework (or library) version.

\item \textsc{Structural coeffects} -- capture additional meta-data related to inputs. This can include
   provenance (source of the input value), usage information (how often is the value
   accessed and in what ways) or security information (whether it contain sensitive data or not).
\end{itemize}
%
This thesis follows the tradition of statically typed programming languages. As such, we
attempt to capture such contextual information in the type system of context-aware programming
languages. The type system should provide both safety guarantees (as in the first three examples)
and also static analysis useful for optimization (as in the last example).

@@

\begin{itemize}
\item In Chapter~\ref{ch:pathways}, we give an overview of related work. Most importantly, we show
  that the idea of context-aware computations can be naturally approached from a number of
  directions developed recently in the theory of programming languages (including type and effect
  systems, categorical semantics and sub-structural logics).

\item In Chapter~\ref{ch:applications}, we present the first contribution of the thesis -- the discovery
  of the connection between a number of existing programming language features that are related
  to context. The chapter presents type systems and semantics for a number of systems and analyses 
  (including data-flow, liveness analysis, distributed programming and Haskell's type classes).
  Our novel presentation reveals their similarity.

### 2014-12-06 text/Chapter02.tex
Proofreading (Chapter 3)

\subsection{Simple coeffect system.}
When writing the judgements of coeffect systems, we want to emphasize the fact that coeffect 
systems talk about \emph{context} rather than \emph{results}. For this reason, we write the 
judgements in the form $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$, associating the additional 
information with the context (left-hand side) of the judgement rather than with the result 
(right-hand side) as in $\Gamma \vdash e : \tau, \cclrd{r}$. This change alone would not be 
very interesting -- we simply used different syntax to write a predicate with four arguments. 
The more interesting difference is how the lambda abstraction rule looks.

>>>>

\subsection{Simple coeffect system.}
\label{sec:path-effects-coeff}
When writing the judgements of coeffect systems, we want to emphasize the fact that coeffect 
systems talk about \emph{context} rather than \emph{results}. For this reason, we write the 
judgements in the form $\coctx{\Gamma}{\cclrd{r}} \vdash e : \tau$, associating the additional 
information with the context (left-hand side) of the judgement rather than with the result 
(right-hand side) as in $\Gamma \vdash e : \tau, \cclrd{r}$. This change alone would not be 
very interesting -- we simply used different syntax to write a predicate with four arguments. 
The more interesting difference is how the lambda abstraction rule looks.
